---
title: "OPIM 5671 Assignment 3 Workings"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---
# Import libraries
```{r}
library(magrittr)
```

# Import data
```{r}
library(readr)
train.df <-
read_csv(
"./Data/train.csv",
col_types = cols(
chris = col_logical(),
customers = col_double(),
date = col_character(),
easter = col_logical(),
open = col_logical(),
ph = col_logical(),
prom = col_logical(),
sch = col_logical()
)
)
summary(train.df)
```
Fix the date column:
```{r}
train.df %<>%
  dplyr::mutate(date = lubridate::ymd(date))
```
Fix the sales column:
```{r}
train.df %<>%
  dplyr::mutate(sales = as.numeric(sales))
summary(train.df)
```
Create the test set:
```{r}
test.df <- train.df %>%
  dplyr::slice(which(is.na(sales)))

summary(test.df)
```
Set the size of the valivation set:
```{r validation size}
valid.size <- 150
```


Crate the validation set:
```{r}
valid.df <- train.df %>%
  dplyr::setdiff(test.df) %>%
  tail(valid.size)

summary(valid.df)  
```
Fix the test training set:
```{r}
train.df %<>%
  dplyr::setdiff(valid.df) %>%
  dplyr::setdiff(test.df)

summary(train.df)
```
# Visualizations
## Data Preparation
Consolidate the training and validation set:
```{r}
consol.df <- train.df %>%
  dplyr::bind_rows(valid.df)

summary(consol.df)
```
Convert the consolidated data frame into a time series object:
```{r }
consol.ts <-
  purrr::map(consol.df[, -1], function(x)
  ts(zoo::zoo(x, order.by = consol.df$date))) %>%
  purrr::invoke(ts.intersect, .)
  
summary(consol.ts)
  
```
## Trend/Seasonality
Plot sales:
```{r}
plot(x = consol.df$date, y = consol.ts[, "sales"], type = "b")
```

Sales by month:
```{r}
consol.df %>%
  dplyr::mutate(month = lubridate::month(date, T),
                year = lubridate::year(date)) %>%
  dplyr::group_by(month, year) %>%
  dplyr::summarise(avg.sales = mean(sales)) %>%
  tidyr::spread(year, avg.sales)
```


## Cross Correlation Function
Plot the ccf of sales against all the other variables:
```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "customers"])
```


```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "open"])
```



```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "prom"])
``` 


```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "sch"])
```

```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "chris"])
```
```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "easter"])
```

```{r}
TSA::prewhiten(consol.ts[, "sales"], consol.ts[, "ph"])
```
## Lagged plots

Plot of sales at various lags:
```{r}
astsa::lag1.plot(consol.ts[,"sales"], 21)
```


# Modelling
## Lagged regression
Prepare the data set:
```{r}
alldata <-
  ts.intersect(
  consol.ts,
  saleslag7 = lag(consol.ts[, "sales"],-7),
  saleslag14 = lag(consol.ts[, "sales"],-14)
  )
  
  dimnames(alldata)[[2]] <- dimnames(alldata)[[2]] %>%
  stringr::str_replace("consol.ts.(.*)", "\\1")
  
  n <- nrow(alldata)
  
  alldata.train <- ts(alldata[1:(n - valid.size),])
  alldata.valid <- ts(alldata[(n - valid.size + 1):n,])
```

Build the model:
```{r}
fit <- lm(log(sales) ~ ., data = alldata.train)
summary(fit)

```
View the residuals:
```{r}
astsa::acf2(residuals(fit))
```
```{r}
plot(fit)
```
Make predictions on the validation set:
```{r}
preds <- predict(fit, alldata.valid) %>%
  exp
  
tgt.cols <- c("date", "sales", "pred.sales", "err.squared")

rest.cols <- setdiff(names(valid.df), tgt.cols)

preds.df <- data.frame(valid.df) %>%
  dplyr::mutate(pred.sales = preds) %>%
  dplyr::mutate(err.squared = (sales - pred.sales) ^ 2) %>%
  dplyr::select(date, sales, pred.sales, err.squared,  match(rest.cols, names(valid.df))) %>%
  dplyr::rename(actual.sales = sales)


```
Compute the RMSE:
```{r}
preds.df %>%
  dplyr::select(err.squared) %>%
  {sqrt(sum(.)/n)}

```

# Feature Enginering

Add log of customers:
```{r}
train.df %>%
  dplyr::mutate(log_customers = log(customers)) ->
  train.df
```
Add month variable:
```{r}
train.df %>%
  dplyr::mutate(month = lubridate::month(date)) ->
  train.df
```
Convert TRUE/FALSE to 0/1:
```{r}
train.df %>%
  dplyr::mutate(
                open = as.numeric(open),
                prom = as.numeric(prom),
                sch = as.numeric(sch),
                chris = as.numeric(chris),
                easter = as.numeric(easter),
                ph = as.numeric(ph)) ->
  train.df
```


Save file:
```{r}
write.csv(train.df, "./Data/train_mod.csv", row.names = F)
```

